Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-13 doc/report/msc-report.tex

It is suitable to to exploit the parallel nature of the records by breaking up
the data sets in pieces and to have each piece be processed separately on a
node in a large compute cluster. It is also suitable to store the data across
many storage nodes in what is called a distributed file system.

@@

OLAP\footnote{On-Line Analytical Processing} is an approach to quickly answer
multi-dimensional analytical queries.

At the core of any OLAP system is the OLAP cube (also known as a
multi-dimensional cube or a hypercube). The cube consists of numeric facts
called measures which are categorized by dimensions. For example, the number
of nails sold is a measure while country, month and type of nail are
dimensions. It is from OLAP that we have appropriated the concept of a data
cube.

@@

A data cube is a conceptual n-dimensional array of values. It aids in
reasoning about dimensions. In the context of this thesis, each dimension of
the cube corresponds to an attribute of the data in question and each point
along the dimension's axis is a value of the corresponding attribute. At any
point where the n dimensions intersect, a measure can be found. For example,
if the dimensions are country and nail type, the point in the cube where
country is equal to Sweden and nail type is equal to tack will contain a
measurement of the units of tack nails sold in Sweden.

@@

Last.fm has a central statistics and metrics tool built on a relational
database. It is possible to import analyzed data to this tool; the data is
presented to the user as graphs. Disadvantages of this system include problems
with new data source additions: they can sometimes not be loaded because they
are to big. If a user wants to perform a custom query, manual intervention by
the developer is necessary. There is no way for this system to handle custom
queries.

@@

Reporting the analyzed data is the final step. This is what the user is
interested in: Viewing the analyzed data. The tool at Last.fm presents the user
with a dashboard that has a selection of graphs on primary metrics; it also has
a list of all available graphs. All the graphs are pre-aggregated, no support
exists for custom queries. Enabling the user to query the aggregated data---for
instance viewing the most popular web browser in a country---would open
possibilities to explore the data. During exploration interesting details could
possibly be discovered; details which may be excluded from the pre-rendered
aggregates.

Since the data sources at Last.fm are time-stamped logs, the constraint to only
handle, time-series was brought upon the Zohmg system. This decision was
reached mainly to keep the implementation simple.

@@

At Last.fm there are several different facilities which produce logs, among
them: web servers, scrobbling\footnote{Last.fm users can submit what music they
listen to, this is called scrobbling.}, radio streaming etc. This data reside
on a commodity hardware cluster used for parallel computation. The system needs
to be able to easily import data from this source.

% storage.
Because the data sources are typically large, the storage need to be efficient.
Storing the analyzed data can be done in different ways, for instance storing
the raw data or just aggregates. Storing the raw data is very expensive, since
it would basically copy the data to the data store. Because of this the decision
was made to only store aggregates.

@@

TODO: How to discuss modelling of data in a general sense, without mentioning
MapReduce and BigTable?

% export.
Due to the several different data sources, there exist several different types
of applications. Therefor the Zohmg system does not implement any user
interface, this is left for the developer to implement.

@@

The four key requirements of the Zohmg system are: ease of use, speed,
scalability, and being able to fit in the existing software and hardware
eco-system.

>>>>

It is suitable to exploit the parallel nature of the records by breaking up
the data sets in pieces and to have each piece be processed on a separate
node in a large compute cluster. It is also suitable to store the data across
many storage nodes in what is called a distributed file system.

@@

OLAP\footnote{On-Line Analytical Processing} is an umbrella term for systems
built in a certain manner and that are optimized quickly answer
multi-dimensional analytical queries.

At the core of any OLAP system is the OLAP cube (also known as a
multi-dimensional cube or a hypercube). The cube consists of numeric facts
-- called measures -- which are grouped by dimensions. For example, the number
of nails sold is a measure while country, month and type of nail are
dimensions. It is from OLAP that [we] have appropriated the concept of a data
cube.

@@

% looking for a mathier explanation here.
A data cube is a conceptual n-dimensional array of values. It aids in
reasoning about dimensions. In the context of this thesis, each dimension of
the cube corresponds to an attribute of the data in question and each point
along the dimension's axis is a value of the corresponding attribute. At any
point where the n dimensions intersect, a measure can be found. For example,
if the dimensions are country and nail type, the point in the cube where
country is equal to Sweden and nail type is equal to tack will contain a
measurement of the units of tack nails sold in Sweden.

@@

Last.fm has a central statistics and metrics tool built on a relational
database. It is possible to import analyzed data to this tool; the data is
presented to the user as graphs. Disadvantages of this system include problems
with new data source additions: they can sometimes not be loaded because they
are too big. If a user wants to perform a custom query, manual intervention by
the developer is necessary. There is no way for this system to handle custom
queries.

@@

Reporting the analyzed data is the final step. This is what the user is
interested in: viewing the analyzed data. The tool at Last.fm presents the user
with a dashboard that has a selection of graphs on primary metrics; it also has
a list of all available graphs. All the graphs are pre-aggregated, no support
exists for custom queries. Enabling the user to query the aggregated data---for
instance viewing the most popular web browser in a country---would open
possibilities to explore the data. During exploration interesting details could
possibly be discovered; details which may be excluded from the pre-rendered
aggregates.

Since the data sources at Last.fm are time-stamped logs, the constraint to only
handle time-series was brought upon the system. This decision was reached
mainly to keep the implementation simple.

@@

At Last.fm there are several different facilities which produce logs, among
them: web servers, scrobbling\footnote{Last.fm users can submit what music they
listen to, this is called scrobbling.}, radio streaming etc. This data reside
on a commodity hardware cluster used for parallel computation. The system needs to
be able to easily import data from these logs, taking in to account both the
locality of the data as well as the structure of the individual data records.

% storage.
Because the data sources are typically large, the storage needs to be efficient.
Storing the analyzed data can be done in different ways, for instance storing
the raw data or just aggregates. Storing the raw data is very expensive, since
it would basically copy the data to the data store. Because of this the decision
was made to only store aggregates.

@@

TODO: How to discuss modelling of data in a general sense, without mentioning
MapReduce and BigTable?
% thinking in simple data structures,
% ..

% export.
Due to the several different data sources, there exist several
different types of applications. Therefor the system does not implement any user
interface other than the REST API\footnote{The REST API is detailed in appendix
X}. Making sense of the data is left for the developer.

@@

The four key requirements of the Zohmg system are: ease of use, speed,
scalability, and adoption, where adoption means being able to fit in the
existing software and hardware eco-system.
% dave probably knows a better word than 'fit'.
